### 2020-02-22 main_text/main_text.tex
Small edits to main text by MJM

	\begin{figure}%[h!]
		\centering
		\includegraphics[width=1\textwidth]{media/Fig1_data.eps}
		\caption{{\bf Experimental data on the \emph{xap} circuit.}
		(A) The expression of the \emph{xapAB} promoter was measured for
		different concentrations of xanthosine in the bacteria's
		environment. The right panel shows the wild type circuit while
		on the left the genes \emph{xapABR} and in the middle the genes
		\emph{xapAB} are removed. The wild type circuit behaves like a
		switch. Note that the fluorescence scales of the middle panel
		and the other two are not comparable.
		(B) The fold-change in gene expression upon addition of
		xanthosine for the wt promoter and for the promoter with only
		the XapR binding site right at the polymerase binding site.
		}
		\label{fig1:data}
	\end{figure}

	When all genes of the \emph{xapABR} circuit are removed, the xanthosine response of our reporter construct disappears (see Fig~\subref{fig1:data}{B}, where the last line is autofluorescence). This is the result of removing the transcription factor XapR that is induced by xanthosine. When only the genes \emph{xapAB} are removed but \emph{xapR} is kept, a response to the xanthosine concentration is regained but it is no longer switch-like (see Fig~\subref{fig1:data}{C}). Instead, the distribution remains unimodal and simply shifts to higher expression as the xanthosine concentration is increased. As we will see later on, this is because the circuit now lacks the positive feedback loop due to the xanthosine membrane transporter XapB.

	Clearly, XapR regulates the \emph{xapAB} promoter. As mentioned in
	the introduction, there are two binding sites for XapR on the
	\emph{xapAB} promoter: one directly at the polymerase binding site
	and one further upstream. When the upstream binding site is removed,
	bimodality is lost and the activation of the \emph{xapAB}-expression
	by xanthosine (via XapR) becomes negligible, as can be seen in
	Fig~\subref{fig1:data}{B}.

	Already, we believe this data, combined with the prior knowledge of the field, clearly suggest a minimal model of the \emph{xapABR} system. The construction and exploration of this model occupies the remainder of this work. We leave it for future work to quantitatively dissect and test the model, in the manner of~\cite{Garcia2011,RazoMejia2018,Chure2019}, for example.

@@

	We consider only the state where both XapR binding sites are occupied as
	active and all other states as inactive, meaning they have transcription
	rate equal to zero. Our experiments show that the expression becomes very weak when one of
	the XapR binding sites is removed from the promoter (see~\nameref{S1_Text}), suggesting that this
	simplification is reasonable. Furthermore, we find that in
	the bistable parameter range, considering the single occupancy states as
	active instead has almost no influence on the results (see
	also~\nameref{S1_Text}).

@@

	\paragraph*{Time evolution and adaptation times.}
	Fig~\subref{fig9:StochT}{A} shows the time evolution of the protein concentration for one typical run of the simulation.
	Again, the simulation was started with a mRNA, protein and intracellular
	xanthosine count of 0 and was run for a simulated time of $5 \cdot 10^5
	\unit{s}$. In this specific example, the fixed point was reached after roughly $1.5 \cdot 10^5
	\unit{s}$ (adaptation time).
	This time varies: Fig~\subref{fig9:StochT}{B} shows the distribution
	of adaptation times from 1000 runs of the simulation (where we
	neglected the around 10 runs with adaptation times larger than $6
	\cdot 10^5 \unit{s}$). In both figures, the trajectory obtained from
	integrating the deterministic system is shown for comparison. It
	clearly agree well with the simulation on average.

	\begin{figure}%[h!]
		\centering
		\includegraphics{media/Fig9_evolution.eps}
		\caption{{\bf Stochastic and deterministic time evolution of protein (XapA/XapB) and adaptation time.}
			(A) Blue shows the results from one run of the stochastic simulation, orange shows the trajectory obtained from solving the deterministic ODE's. The simulation was run once for $5 \cdot 10^5 \unit{s}$ and started
			at an mRNA, protein and intracellular xanthosine count of 0. The parameters that were used are the same as in
			Table~\ref{table1:nondim}, the only exception being the
			extracellular xanthosine concentration, which was chosen to be
			$\n{[c]_a} = 25$ (recall $\n{[c]_{a}} \defeq \frac{c}{K_{\n{a}}}$,
			$K_{\n{a}} = 5 \cdot 10^{-5} \unit{M}$) just as in
			Fig~\subref{fig8:stochC}{(C)}. (B) Blue shows the time until 90\% of the protein concentration at the upper fixed point was reached in 1000 runs of the simulation (same conditions as (A)). The orange line shows the corresponding deterministic time. The green dashed line right next to it indicated the mean of the blue distribution.}
		\label{fig9:StochT}
	\end{figure}

>>>>

	\begin{figure}%[h!]
		\centering
		\includegraphics[width=1\textwidth]{media/Fig1_data.eps}
		\caption{{\bf Experimental data on the \emph{xap} circuit.}
		(A) The expression of the \emph{xapAB} promoter was measured for
		different extracellular concentrations of xanthosine.
		The right panel shows the wild type circuit while
		on the left the genes \emph{xapABR} and in the middle the genes
		\emph{xapAB} are removed. The wild type circuit behaves like a
		switch. Note that the fluorescence scale of the middle panel
		is not comparable with the other two,
		and also that the chosen xanthosine concentrations are different.
		(B) The fold-change in gene expression upon addition of
		xanthosine for the wt promoter (left panel)
		and for the promoter with only the XapR binding site adjacent
		to the polymerase binding site (right panel).
		Note that the fold-change used here differs from fold-change in,
		e.g.,~\cite{Garcia2011,RazoMejia2018,Chure2019},
		in that no subtraction of autofluorescence was carried out,
		which is adequate for the qualitative comparison
		of these two promoters.
		}
		\label{fig1:data}
	\end{figure}

	When all genes of the \emph{xapABR} circuit are removed, the
	xanthosine response of our reporter construct disappears (see
	Fig~\subref{fig1:data}{A}, left panel).
	This is the result of removing the transcription factor XapR that is
	induced by xanthosine. When only the genes \emph{xapAB} are removed
	but \emph{xapR} is kept, a response to the xanthosine concentration
	is regained but it is no longer switch-like (see
	Fig~\subref{fig1:data}{A}, middle panel). Instead, the distribution remains
	unimodal and simply shifts to higher expression as the xanthosine
	concentration is increased. As we will see later on, this is because
	the circuit now lacks the positive feedback loop due to the
	xanthosine membrane transporter XapB.

	Can we gain further insight into how XapR regulates
	the \emph{xapAB} promoter? As mentioned in
	the introduction, there are two binding sites for XapR on the
	\emph{xapAB} promoter: one directly at the polymerase binding site
	and one further upstream.
	Working in the \textDelta\emph{xapAB} background,
	we measured the expression level of our reporter when driven by two
	constructs: the wild-type \emph{xapAB} promoter
	(Fig~\subref{fig1:data}{B}, left panel), and the \emph{xapAB}
	promoter with the upstream binding site removed
	(Fig~\subref{fig1:data}{B}, right panel).
	Clearly, removing one XapR binding site dramatically reduced the
	responsiveness of the promoter to xanthosine (though not apparent
	from Fig~\subref{fig1:data}{B}, this response remains detectable).

	Already, we believe these data, combined with the prior knowledge of
	the field, clearly suggest a minimal model of the \emph{xapABR}
	system. The construction and exploration of this model occupies the
	remainder of this work. We leave it for future work to
	quantitatively dissect and test the model, in the manner
	of~\cite{Garcia2011,RazoMejia2018,Chure2019}, for example.

@@

	We consider only the state where both XapR binding sites are occupied as
	active and all other states as inactive, meaning they have transcription
	rate equal to zero. Our experiments show that the expression becomes very weak when one of
	the XapR binding sites is removed from the promoter (see
	Fig~\subref{fig1:data}{B} and~\nameref{S1_Text}), suggesting that
	this simplification is reasonable. Furthermore, we find that in
	the bistable parameter range, considering the single occupancy states as
	active instead has almost no influence on the results (see
	also~\nameref{S1_Text}).

@@

	\paragraph*{Time evolution and adaptation times.}
	Fig~\subref{fig9:StochT}{A} shows the time evolution of the protein concentration for one typical run of the simulation.
	Again, the simulation was started with a mRNA, protein and intracellular
	xanthosine count of 0 and was run for a simulated time of $5 \cdot 10^5
	\unit{s}$. In this specific example, the fixed point was reached after roughly $1.5 \cdot 10^5
	\unit{s}$ (adaptation time).
	This time varies: Fig~\subref{fig9:StochT}{B} shows the distribution
	of adaptation times from 1000 runs of the simulation.
	In both figures, the trajectory obtained from
	integrating the deterministic system is shown for comparison. It
	clearly agrees well with the simulation on average.

	\begin{figure}%[h!]
		\centering
		\includegraphics{media/Fig9_evolution.eps}
		\caption{{\bf Stochastic and deterministic time evolution of protein (XapA/XapB) and adaptation time.}
			(A) Blue shows the results from one run of the stochastic simulation, orange shows the trajectory obtained from solving the deterministic ODE's. The simulation was run once for $5 \cdot 10^5 \unit{s}$ and started
			at an mRNA, protein and intracellular xanthosine count of 0. The parameters that were used are the same as in
			Table~\ref{table1:nondim}, the only exception being the
			extracellular xanthosine concentration, which was chosen to be
			$\n{[c]_a} = 25$ (recall $\n{[c]_{a}} \defeq \frac{c}{K_{\n{a}}}$,
			$K_{\n{a}} = 5 \cdot 10^{-5} \unit{M}$) just as in
			Fig~\subref{fig8:stochC}{(C)}.
			(B) In blue we plot the waiting time distribution for 1000
			runs of the simulation (same conditions as (A)) to reach
			the steady-state mean.
			We define this time as the elapsed time when the protein
			copy number first reaches 90\% of its value at the upper
			fixed point.
			To better visualize the bulk of the distribution, we
			excluded from the plot $\sim10$ runs with adaptation times
			larger than $6 \cdot 10^5 \unit{s}$.
			The green dashed line indicates the mean of the blue distribution.
			The orange line shows the corresponding deterministic time.
			}
		\label{fig9:StochT}
	\end{figure}
